Learning to Infer Graphics Programs from Hand-Drawn Images

### 2017-10-27 short.tex
final version before I do some final proofreading

We present a hybrid architecture for inferring graphics programs that is structured around these two steps. For the first step, our approach uses a deep network to infer a set of primitive shape-drawing commands. We refer to this set as a \emph{trace set} since it corresponds to the set of commands in a program's execution trace but lacks the high-level structure that determines how the program decided to issue them.
%For added robustness, we train the network to produce a proposal distribution of the most likely traces and use stochastic search to find the trace that best matches the image.
%The network is trained from an automatically-generated corpus of synthetic image-generating programs.
The second step involves synthesizing a high-level program capable of producing the trace set identified by the first phase.
%In principle, we could also train a neural network to learn this mapping. In practice, though, this would be challenging because it would no longer be sufficient to train the network with a synthetic corpus of randomly generated programs, since our goal is to generate programs that satisfy certain semantic constraints that ensure that we get the program that the user most likely intended.
%This paper argues that this stage is best
We achieve this by \emph{constraint-based program synthesis} \citep{solar2008program}. The program synthesizer searches the space of possible programs for one capable of producing the desired trace -- inducing structures like
symmetries, loops, or conditional branches.
Although these program synthesizers do not need any training data,
we show how to learn a \emph{search policy} 
in order to synthesize programs an order of magnitude faster than constraint-based synthesis techniques alone.

@@

We developed a deep network architecture for efficiently inferring a
trace set, $T$, from an image, $I$.
Our model is a combination of  ideas from 
Neurally-Guided Procedural Models~\citep{ritchie2016neurally}
and Attend-Infer-Repeat~\citep{eslami1603attend}.
Our model constructs the
trace one drawing command at a time.  When predicting the next drawing
command, the network takes as input the target image $I$ as well as
the rendered output of previous drawing commands.  Intuitively, the
network looks at the image it wants to explain, as well as what it has
already drawn.  It then decides either to stop drawing or proposes
another drawing command to add to the trace; if it decides
to continue drawing, the predicted primitive is rendered to its
``canvas'' and the process repeats.

Fig.~\ref{architecture} illustrates this architecture.  We first
pass a $256\times 256$ target image and a rendering of the trace so
far (encoded as a two-channel image) to a convolutional network. Given
the features extracted by the convnet, a multilayer perceptron then
predicts a distribution over the next drawing command to add to the
trace; see Tbl.~\ref{drawingCommandTable} for the drawing commands we currently model.  We predict the drawing
command token-by-token,\footnote{Tokens are atomic symbols in the programming language  -- so the model is \emph{not} predicting lines of code character-by-character, but instead the model knows about the syntactic structure of primitive drawing commands} conditioning each token both on the image
features and on the previously generated tokens. We also use a
differentiable attention mechanism (Spatial Transformer
Networks:~\cite{jaderberg2015spatial}) to let the model attend to
different regions of the image while predicting each token. In Fig.~\ref{architecture} the network first decided to emit the \verb|circle| token
conditioned on the image features, then it emitted the $x$ coordinate of
the circle conditioned on the image features and the \verb|circle|
token, and finally it predicted the $y$ coordinate of the circle
conditioned on the image features, the \verb|circle| token, and the
$x$ coordinate. We currently constrain
coordinates to lie on a discrete $16\times 16$ grid,
but the grid could be made arbitrarily fine.  See supplement for the full details
of the architecture, which we implemented in
Tensorflow~\citep{tensorflow2015-whitepaper}.
%% \remark{There are some more details that are important to provide
%%   about this architecture in the supplement: the
%%   functional form(s) of the probability distributions over tokens, the
%%   network layer sizes, which MLPs share parameters, etc.}

@@

Fig.~\ref{syntheticResults}
compares the neural network with SMC against the neural network by
itself or SMC by itself.  Only the combination of the two passes a
critical test of generalization: when trained on images with $\leq 12$
objects, it successfully parses scenes with many more objects than the
training data.
We compare with a baseline that models the problem as image captioning (LSTM in Fig.~\ref{syntheticResults}).
Given the target image, this baseline produces the trace in one shot by
using a CNN to extract features of the input which are passed to an LSTM which finally predicts
the trace token-by-token.
This architecture is used in several successful neural models of image captioning (e.g.,~\cite{vinyals2015show}),
but, for this domain, does not suffice to parse cluttered scenes with many objects.

@@

Our neurally-guided SMC procedure
used pixel-wise distance as a surrogate for a likelihood function ($L(\cdot|\cdot)$ in section~\ref{neuralNetworkSection}).
But pixel-wise distance fares poorly on hand drawings, which never exactly match
the model's renders.
So, for hand drawings,
we \emph{learn} a surrogate likelihood function,
$L_{\text{learned}}(\cdot|\cdot)$.
The density $L_{\text{learned}}(\cdot|\cdot)$ is predicted by a convolutional network that we train to predict
the distance between two traces conditioned upon their renderings.
We train our likelihood surrogate to approximate the symmetric difference,
which is  the number of drawing commands by which two traces differ:
  \begin{equation}
    -\log L_{\text{learned}}(\text{render}(T_1)|\text{render}(T_2))\approx |T_1 - T_2| + |T_2 - T_1|\label{symmetricDistance}
    \end{equation}

@@

\begin{figure}[h]\centering
  \begin{minipage}[c]{0.57\textwidth} 
    \centering  \includegraphics[width = 8cm]{figures/drawingAccuracy.png}
  \end{minipage}\hfill%
      \begin{minipage}[c]{0.4\textwidth} 
  \caption{How close are the model's outputs to the ground truth on hand drawings, as we consider larger sets of samples (1, 5, 100)?
    Distance to ground truth trace measured by the intersection over union of predicted vs. ground truth traces. (a) for 63\% of drawings the model's top prediction is exactly correct; (b) for 70\% of drawings the ground truth is in the top 5 model predictions; (c) for 4\% of drawings all of the model outputs have no overlap with the ground truth trace. Red: the full model. Other colors: lesioned versions of our model.}\label{drawingIntersectionOverUnion}
      \end{minipage}
            \vspace{-1cm}
\end{figure}

@@

\section{Synthesizing graphics programs from trace sets}\label{programSynthesisSection}
Although the trace set of a graphics program describes the contents
of a scene, it does not encode higher-level features of the image,
such as repeated motifs or symmetries.  A \emph{graphics
  program} better describes such structures.
We seek to synthesize graphics programs from their traces.

@@

Given the DSL and a trace $T$, we want to recover a program that both evaluates to $T$
and, at the same time, is the ``best'' explanation of $T$.
For example, we might prefer more general programs or, in the spirit of Occam's razor,
prefer shorter programs.
We wrap these intuitions up into a cost function over programs,
and seek the minimum cost program consistent with $T$:
\begin{equation}
  \text{program}(T) = \argmin_{p\in \text{DSL, s.t. }p \text{ evaluates to } T} \text{cost}(p)\label{programObjective}
\end{equation}
We define the
cost of a program to be the number of Statement's it contains (Tbl.~\ref{DSL}).
We also penalize using many different numerical constants; see supplement.
%\remark{The flow here is a bit off/backwards. ``We want a program that evaluates to $T$ and also minimizes some measure of program cost''---why do we care about cost? It'd be better to start by making a ``Bayesian Occam's razor'' appeal (e.g. the most compact/general program is the more likely explanation) and then saying that one way to do this is to minimize a cost function which is proportional to program length.}

The constrained optimization problem in
Eq.~\ref{programObjective} is intractable in general, but there
exist efficient-in-practice tools for finding exact solutions to such
program synthesis problems. We use the state-of-the-art Sketch
tool~\cite{solar2008program}.
%At a high level,
Sketch takes as input a space of programs, along with
a specification of the program's behavior and optionally a cost
function.  It translates the synthesis problem into a constraint
satisfaction problem and then uses a SAT solver to find a minimum-cost
program satisfying the specification.  Sketch requires a
 \emph{finite program space}, which here means that the depth of the
program syntax tree is bounded (we set the bound to 3),
but has the guarantee that it 
always eventually finds the globally optimal solution.
In exchange for this optimality guarantee
it comes with no guarantees
on runtime.
For our domain synthesis times vary from minutes to hours,
with 27\% of the drawings timing out the synthesizer after 1 hour.
Fig.~\ref{exampleSynthesisResults} shows programs recovered by our system.
A main impediment to our use of these general techniques is
the prohibitively high cost of searching for programs.
We next describe how to learn to synthesize programs much faster (Sec.~\ref{learningASearchPolicy}),
timing out on 2\% of the drawings and solving 58\% of problems within a minute.

@@

\begin{figure}[t]
%  \vspace{-0.6cm}
\centering  \begin{tabular}{m{1.5cm}llc}
    \toprule
    \textbf{Drawing}&\textbf{Trace Set}&\textbf{Program}&%\begin{tabular}{c}
      \textbf{Compression factor}%\\\textbf{factor}
      %      \end{tabular}\\
      \\
    \midrule
    \includegraphics[width = \exampleDrawingSize]{figures/expert-29-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Line(2,15, 4,15)
Line(4,9, 4,13)
Line(3,11, 3,14)
Line(2,13, 2,15)
Line(3,14, 6,14)
Line(4,13, 8,13)
\end{lstlisting}
\end{minipage}&     \begin{minipage}{\exampleProgramSize} \begin{lstlisting}
for(i<3)
 line(i,-1*i+6,
      2*i+2,-1*i+6)
 line(i,-2*i+4,i,-1*i+6)
       \end{lstlisting}
     \end{minipage}&$\frac{6}{3} = 2\text{x}$\\\midrule
     \includegraphics[width = \exampleDrawingSize]{figures/expert-52-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Line(5,13,2,10,arrow)
Circle(5,9)
Circle(8,5)
Line(2,8, 2,6,arrow)
Circle(2,5)
\end{lstlisting}
\small\emph{... etc. ...; 13 lines}
\end{minipage}&
             \begin{minipage}{\exampleProgramSize}\begin{lstlisting}
circle(4,10)
for(i<3)
 circle(-3*i+7,5)
 circle(-3*i+7,1)
 line(-3*i+7,4,-3*i+7,2,arrow)
 line(4,9,-3*i+7,6,arrow)
\end{lstlisting}
\end{minipage}&$\frac{13}{6} = 2.2\text{x}$\\\midrule
    \includegraphics[width = \exampleDrawingSize]{figures/expert-38-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Circle(5,8)
Circle(2,8)
Circle(8,11)
Line(2,9, 2,10)
Circle(8,8)
Line(3,8, 4,8)
Line(3,11, 4,11)
\end{lstlisting}
  \small\emph{... etc. ...; 21 lines}
\end{minipage}&\begin{minipage}{\exampleProgramSize}
\begin{lstlisting}
for(i<3)
 for(j<3)
  if(j>0)
   line(-3*j+8,-3*i+7,
        -3*j+9,-3*i+7)
   line(-3*i+7,-3*j+8,
        -3*i+7,-3*j+9)
  circle(-3*j+7,-3*i+7)
\end{lstlisting}
\end{minipage}&$\frac{21}{6} = 3.5\text{x}$\\\midrule
\includegraphics[width = \exampleDrawingSize]{figures/expert-101-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Rectangle(1,10,3,11)
Rectangle(1,12,3,13)
Rectangle(4,8,6,9)
Rectangle(4,10,6,11)
\end{lstlisting}
  \small\emph{... etc. ...; 16 lines}
\end{minipage}&\begin{minipage}{\exampleProgramSize}
\begin{lstlisting}
for(i<4)
 for(j<4)
  rectangle(-3*i+9,-2*j+6,
            -3*i+11,-2*j+7)
\end{lstlisting}
\end{minipage}&$\frac{16}{3} = 5.3\text{x}$\\\midrule
\includegraphics[width = \exampleDrawingSize]{figures/expert-75-extra-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Line(11,14,13,14,arrow)
Circle(10,10)
Line(10,13,10,11,arrow)
Circle(6,10)
\end{lstlisting}
    \small\emph{... etc. ...; 15 lines}
\end{minipage}&
\begin{minipage}{\exampleProgramSize}
    \begin{lstlisting}
for(i<4)
 line(-4*i+13,4,-4*i+13,2,arrow)
 for(j<3)
  if(j>0)
   circle(-4*i+13,4*j+-3)
  line(-4*j+10,5,-4*j+12,5,
       arrow)
\end{lstlisting}
  \end{minipage}&$\frac{15}{6} = 2.5\text{x}$\\\midrule

@@

\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Circle(2,8)
Rectangle(6,9, 7,10)
Circle(8,8)
Rectangle(6,12, 7,13)
Rectangle(3,9, 4,10)
\end{lstlisting}\small\emph{... etc. ...; 9 lines}
  \end{minipage}&\begin{minipage}{\exampleProgramSize}
\begin{lstlisting}
reflect(y=8)
 for(i<3)
  if(i>0)
   rectangle(3*i-1,2,3*i,3)
  circle(3*i+1,3*i+1)
\end{lstlisting}
\end{minipage}&$\frac{9}{5} = 1.8\text{x}$ \\\bottomrule
  \end{tabular}
\caption{Example drawings (left), their ground truth trace sets (middle left), and programs synthesized from these traces (middle right). Compared to the trace the programs are more compressive (right: programs have fewer lines than traces) and automatically group together related drawing commands. Note the nested loops in the Ising model, special case conditionals for the HMM, combination of symmetry and iteration in the bottom figure,  affine transformations in the top figure, and the complicated program in the second figure to bottom.}\label{exampleSynthesisResults}
  \end{figure}

@@

\begin{table}[h]\centering
  \caption{Parameterization of different ways of posing the program synthesis problem. The policy learns to choose parameters likely to quickly yield a minimal cost program. We slightly abuse notation by writing $\sigma $ to mean an assignment to each of these parameters (so $\sigma $ assumes one of 24 different values) and to mean the set of programs selected by that parameterization (so $\sigma \subseteq \text{DSL}$)}\label{policyOutput}
  \begin{tabular}{lll}\toprule
  Parameter&Description&Range\\\midrule
  Loops?&Is the program allowed to loop?&$\{\text{True},\text{False}\}$\\
  Reflects?&Is the program allowed to have reflections?&$\{\text{True},\text{False}\}$\\
  Incremental?&Solve the problem piece-by-piece or all at once?&$\{\text{True},\text{False}\}$\\
  Maximum depth& Bound on the depth of the program syntax tree&$\{1,2,3\}$
  \\\bottomrule
  \end{tabular}
  \end{table}
Given a search policy, what is the best way of using it to quickly find minimum cost programs?
We use a \emph{bias-optimal search algorithm}~\citep{schmidhuber2004optimal}:

@@

Intuitively, an $n$-bias optimal search algorithm
spends at least $\probability_{\text{bias}}[\sigma ]/n$ of its time
exploring solution space $\sigma $.
An example of a $1$-bias optimal search algorithm is a time-sharing system
that allocates $\probability_{\text{bias}}[\sigma ]$ of its time to trying $\sigma $.
We construct a $1$-bias optimal search algorithm for our domain by
identifying $\probability_{\text{bias}}[\sigma ] = \pi_\theta(\sigma |T)$
and $t(\sigma )  $ with how long the synthesizer takes to search  the programs in $\sigma $.
Now in theory any $\pi_\theta(\cdot |\cdot ) $ is a bias-optimal searcher.
But the actual runtime of the algorithm depends strongly upon
the bias $\probability_{\text{bias}}[\cdot ]$.
Our new approach is to learn $\probability_{\text{bias}}[\cdot ]$
by picking the policy minimizing the
expected bias-optimal time to solve a training corpus, $\mathcal{D}$, of graphics program synthesis problems:
\begin{align}
\textsc{Loss}(\theta ; \mathcal{D})& =  \expect_{T\sim\mathcal{D}}\left[ \min_{\sigma\in \text{\textsc{Best}}(T)}\frac{t(\sigma | T)}{\pi_\theta (\sigma | T)}\right] + \lambda \Vert\theta\Vert_2^2\label{policyLoss}\\
\text{where }  \sigma \in\text{\textsc{Best}}(T) &\text{ if  a minimum cost program for }T \text{ is in }\sigma .\nonumber %\text{cost}(\text{program}(p)) = \min_{p\in \sigma \text{, s.t. }p\text{ evaluates to }T} \text{cost}(p) \nonumber
\end{align}
To generate a training corpus for learning a policy which minimizes this loss,
we synthesized minimum cost programs for each trace of our hand drawings
and for each $\sigma $.
%Solving these
%\footnote{There are 100 the drawings and $2\times 2\times 2\times 3 = 24$ different values of $\sigma $}
%synthesis problems takes a little over a day on 40 CPUs.
%With the training set in hand we  %This loss is differentiable but nonconvex even if $\pi_{\theta}(\cdot |\cdot )^{-1}$ is convex.
We locally minimize this loss using gradient descent. % while annealing a softened minimum to the hard minimization equation~\ref{policyLoss}.
See supplement for details.
Because we want to learn a policy from only $100$ hand-drawn sketches,
we chose a simple low-capacity, bilinear model for a policy:
\begin{equation}
  \pi_{\theta}(\sigma |T)\propto \exp \left( \phi_{\text{params}}(\sigma )^\top\theta \phi_{\text{trace}}(T)\right)
\end{equation}
where $\phi_{\text{params}}(\sigma  )$ is a one-hot encoding of
the parameter settings of $\sigma $ (see Tbl.~\ref{policyOutput})
and $\phi_{\text{trace}}(T)$ extracts a few simple features of the trace $T$;
see supplement for details.

>>>>

We present a hybrid architecture for inferring graphics programs that is structured around these two steps. For the first step, our approach uses a deep network to infer a set of primitive shape-drawing commands. We refer to this set as a \emph{trace set} since it corresponds to the set of commands in a program's execution trace but lacks the high-level structure that determines how the program decided to issue them.
%For added robustness, we train the network to produce a proposal distribution of the most likely traces and use stochastic search to find the trace that best matches the image.
%The network is trained from an automatically-generated corpus of synthetic image-generating programs.
The second step involves synthesizing a high-level program capable of producing the trace set identified by the first phase.
%In principle, we could also train a neural network to learn this mapping. In practice, though, this would be challenging because it would no longer be sufficient to train the network with a synthetic corpus of randomly generated programs, since our goal is to generate programs that satisfy certain semantic constraints that ensure that we get the program that the user most likely intended.
%This paper argues that this stage is best
We achieve this by \emph{constraint-based program synthesis} \citep{solar2008program}. The program synthesizer searches the space of possible programs for one capable of producing the desired trace set -- inducing structures like
symmetries, loops, or conditional branches.
Although these program synthesizers do not need any training data,
we show how to learn a \emph{search policy} 
in order to synthesize programs an order of magnitude faster than constraint-based synthesis techniques alone.

@@

We developed a deep network architecture for efficiently inferring a
trace set, $T$, from an image, $I$.
Our model is a combination of  ideas from 
Neurally-Guided Procedural Models~\citep{ritchie2016neurally}
and Attend-Infer-Repeat~\citep{eslami1603attend}.
Our model constructs the
trace one drawing command at a time.  When predicting the next drawing
command, the network takes as input the target image $I$ as well as
the rendered output of previous drawing commands.  Intuitively, the
network looks at the image it wants to explain, as well as what it has
already drawn.  It then decides either to stop drawing or proposes
another drawing command to add to the trace set; if it decides
to continue drawing, the predicted primitive is rendered to its
``canvas'' and the process repeats.

Fig.~\ref{architecture} illustrates this architecture.  We first
pass a $256\times 256$ target image and a rendering of the trace so
far (encoded as a two-channel image) to a convolutional network. Given
the features extracted by the convnet, a multilayer perceptron then
predicts a distribution over the next drawing command to add to the
trace set; see Tbl.~\ref{drawingCommandTable} for the drawing commands we currently model.  We predict the drawing
command token-by-token,\footnote{Tokens are atomic symbols in the programming language  -- so the model is \emph{not} predicting lines of code character-by-character, but instead the model knows about the syntactic structure of primitive drawing commands} conditioning each token both on the image
features and on the previously generated tokens. We also use a
differentiable attention mechanism (Spatial Transformer
Networks:~\cite{jaderberg2015spatial}) to let the model attend to
different regions of the image while predicting each token. In Fig.~\ref{architecture} the network first decided to emit the \verb|circle| token
conditioned on the image features, then it emitted the $x$ coordinate of
the circle conditioned on the image features and the \verb|circle|
token, and finally it predicted the $y$ coordinate of the circle
conditioned on the image features, the \verb|circle| token, and the
$x$ coordinate. We currently constrain
coordinates to lie on a discrete $16\times 16$ grid,
but the grid could be made arbitrarily fine.  See supplement for the full details
of the architecture, which we implemented in
Tensorflow~\citep{tensorflow2015-whitepaper}.
%% \remark{There are some more details that are important to provide
%%   about this architecture in the supplement: the
%%   functional form(s) of the probability distributions over tokens, the
%%   network layer sizes, which MLPs share parameters, etc.}

@@

Fig.~\ref{syntheticResults}
compares the neural network with SMC against the neural network by
itself or SMC by itself.  Only the combination of the two passes a
critical test of generalization: when trained on images with $\leq 12$
objects, it successfully parses scenes with many more objects than the
training data.
We compare with a baseline that models the problem as image captioning (LSTM in Fig.~\ref{syntheticResults}).
Given the target image, this baseline produces the trace set in one shot by
using a CNN to extract features of the input which are passed to an LSTM which finally predicts
the trace token-by-token.
This architecture is used in several successful neural models of image captioning (e.g.,~\cite{vinyals2015show}),
but, for this domain, does not suffice to parse cluttered scenes with many objects.

@@

Our neurally-guided SMC procedure
used pixel-wise distance as a surrogate for a likelihood function ($L(\cdot|\cdot)$ in section~\ref{neuralNetworkSection}).
But pixel-wise distance fares poorly on hand drawings, which never exactly match
the model's renders.
So, for hand drawings,
we \emph{learn} a surrogate likelihood function,
$L_{\text{learned}}(\cdot|\cdot)$.
The density $L_{\text{learned}}(\cdot|\cdot)$ is predicted by a convolutional network that we train to predict
the distance between two trace sets conditioned upon their renderings.
We train our likelihood surrogate to approximate the symmetric difference,
which is  the number of drawing commands by which two trace sets  differ:
  \begin{equation}
    -\log L_{\text{learned}}(\text{render}(T_1)|\text{render}(T_2))\approx |T_1 - T_2| + |T_2 - T_1|\label{symmetricDistance}
    \end{equation}

@@

\begin{figure}[h]\centering
  \begin{minipage}[c]{0.57\textwidth} 
    \centering  \includegraphics[width = 8cm]{figures/drawingAccuracy.png}
  \end{minipage}\hfill%
      \begin{minipage}[c]{0.4\textwidth} 
  \caption{How close are the model's outputs to the ground truth on hand drawings, as we consider larger sets of samples (1, 5, 100)?
    Distance to ground truth trace set measured by the intersection over union of predicted vs. ground truth. (a) for 63\% of drawings the model's top prediction is exactly correct; (b) for 70\% of drawings the ground truth is in the top 5 model predictions; (c) for 4\% of drawings all of the model outputs have no overlap with the ground truth trace. Red: the full model. Other colors: lesioned versions of our model.}\label{drawingIntersectionOverUnion}
      \end{minipage}
            \vspace{-1cm}
\end{figure}

@@

\section{Synthesizing graphics programs from trace sets}\label{programSynthesisSection}
Although the trace set of a graphics program describes the contents
of a scene, it does not encode higher-level features of the image,
such as repeated motifs or symmetries.  A \emph{graphics
  program} better describes such structures.
We seek to synthesize graphics programs from their trace sets.

@@

Given the DSL and a trace set $T$, we want to recover a program that both evaluates to $T$
and, at the same time, is the ``best'' explanation of $T$.
For example, we might prefer more general programs or, in the spirit of Occam's razor,
prefer shorter programs.
We wrap these intuitions up into a cost function over programs,
and seek the minimum cost program consistent with $T$:
\begin{equation}
  \text{program}(T) = \argmin_{p\in \text{DSL, s.t. }p \text{ evaluates to } T} \text{cost}(p)\label{programObjective}
\end{equation}
We define the
cost of a program to be the number of Statement's it contains (Tbl.~\ref{DSL}).
We also penalize using many different numerical constants; see supplement.
%\remark{The flow here is a bit off/backwards. ``We want a program that evaluates to $T$ and also minimizes some measure of program cost''---why do we care about cost? It'd be better to start by making a ``Bayesian Occam's razor'' appeal (e.g. the most compact/general program is the more likely explanation) and then saying that one way to do this is to minimize a cost function which is proportional to program length.}

The constrained optimization problem in
Eq.~\ref{programObjective} is intractable in general, but there
exist efficient-in-practice tools for finding exact solutions to such
program synthesis problems. We use the state-of-the-art Sketch
tool~\cite{solar2008program}.
%At a high level,
Sketch takes as input a space of programs, along with
a specification of the program's behavior and optionally a cost
function.  It translates the synthesis problem into a constraint
satisfaction problem and then uses a SAT solver to find a minimum-cost
program satisfying the specification.  Sketch requires a
 \emph{finite program space}, which here means that the depth of the
program syntax tree is bounded (we set the bound to 3),
but has the guarantee that it 
always eventually finds the globally optimal solution.
In exchange for this optimality guarantee
it comes with no guarantees
on runtime.
For our domain synthesis times vary from minutes to hours,
with 27\% of the drawings timing out the synthesizer after 1 hour.
Tbl.~\ref{exampleSynthesisResults} shows programs recovered by our system.
A main impediment to our use of these general techniques is
the prohibitively high cost of searching for programs.
We next describe how to learn to synthesize programs much faster (Sec.~\ref{learningASearchPolicy}),
timing out on 2\% of the drawings and solving 58\% of problems within a minute.

@@

\begin{table}[t]
  \caption{Example drawings (left), their ground truth trace sets (middle left), and programs synthesized from these traces (middle right). Compared to the trace the programs are more compressive (right: programs have fewer lines than traces) and automatically group together related drawing commands. Note the nested loops in the Ising model, special case conditionals for the HMM, combination of symmetry and iteration in the bottom figure,  affine transformations in the top figure, and the complicated program in the second figure to bottom.}\label{exampleSynthesisResults}
%  \vspace{-0.6cm}
\centering  \begin{tabular}{m{1.5cm}llc}
    \toprule
    \textbf{Drawing}&\textbf{Trace Set}&\textbf{Program}&%\begin{tabular}{c}
      \textbf{Compression factor}%\\\textbf{factor}
      %      \end{tabular}\\
      \\
    \midrule
    \includegraphics[width = \exampleDrawingSize]{figures/expert-29-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Line(2,15, 4,15)
Line(4,9, 4,13)
Line(3,11, 3,14)
Line(2,13, 2,15)
Line(3,14, 6,14)
Line(4,13, 8,13)
\end{lstlisting}
\end{minipage}&     \begin{minipage}{\exampleProgramSize} \begin{lstlisting}
for(i<3)
 line(i,-1*i+6,
      2*i+2,-1*i+6)
 line(i,-2*i+4,i,-1*i+6)
       \end{lstlisting}
     \end{minipage}&$\frac{6}{3} = 2\text{x}$\\\midrule
     \includegraphics[width = \exampleDrawingSize]{figures/expert-52-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Line(5,13,2,10,arrow)
Circle(5,9)
Circle(8,5)
Line(2,8, 2,6,arrow)
Circle(2,5)
\end{lstlisting}
\small\emph{... etc. ...; 13 lines}
\end{minipage}&
             \begin{minipage}{\exampleProgramSize}\begin{lstlisting}
circle(4,10)
for(i<3)
 circle(-3*i+7,5)
 circle(-3*i+7,1)
 line(-3*i+7,4,-3*i+7,2,arrow)
 line(4,9,-3*i+7,6,arrow)
\end{lstlisting}
\end{minipage}&$\frac{13}{6} = 2.2\text{x}$\\\midrule
    \includegraphics[width = \exampleDrawingSize]{figures/expert-38-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Circle(5,8)
Circle(2,8)
Circle(8,11)
Line(2,9, 2,10)
Circle(8,8)
Line(3,8, 4,8)
Line(3,11, 4,11)
\end{lstlisting}
  \small\emph{... etc. ...; 21 lines}
\end{minipage}&\begin{minipage}{\exampleProgramSize}
\begin{lstlisting}
for(i<3)
 for(j<3)
  if(j>0)
   line(-3*j+8,-3*i+7,
        -3*j+9,-3*i+7)
   line(-3*i+7,-3*j+8,
        -3*i+7,-3*j+9)
  circle(-3*j+7,-3*i+7)
\end{lstlisting}
\end{minipage}&$\frac{21}{6} = 3.5\text{x}$\\\midrule
\includegraphics[width = \exampleDrawingSize]{figures/expert-101-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Rectangle(1,10,3,11)
Rectangle(1,12,3,13)
Rectangle(4,8,6,9)
Rectangle(4,10,6,11)
\end{lstlisting}
  \small\emph{... etc. ...; 16 lines}
\end{minipage}&\begin{minipage}{\exampleProgramSize}
\begin{lstlisting}
for(i<4)
 for(j<4)
  rectangle(-3*i+9,-2*j+6,
            -3*i+11,-2*j+7)
\end{lstlisting}
\end{minipage}&$\frac{16}{3} = 5.3\text{x}$\\\midrule
\includegraphics[width = \exampleDrawingSize]{figures/expert-75-extra-trim.png}&
\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Line(11,14,13,14,arrow)
Circle(10,10)
Line(10,13,10,11,arrow)
Circle(6,10)
\end{lstlisting}
    \small\emph{... etc. ...; 15 lines}
\end{minipage}&
\begin{minipage}{\exampleProgramSize}
    \begin{lstlisting}
for(i<4)
 line(-4*i+13,4,-4*i+13,2,arrow)
 for(j<3)
  if(j>0)
   circle(-4*i+13,4*j+-3)
  line(-4*j+10,5,-4*j+12,5,
       arrow)
\end{lstlisting}
  \end{minipage}&$\frac{15}{6} = 2.5\text{x}$\\\midrule

@@

\begin{minipage}{\exampleTraceSize}\begin{lstlisting}
Circle(2,8)
Rectangle(6,9, 7,10)
Circle(8,8)
Rectangle(6,12, 7,13)
Rectangle(3,9, 4,10)
\end{lstlisting}\small\emph{... etc. ...; 9 lines}
  \end{minipage}&\begin{minipage}{\exampleProgramSize}
\begin{lstlisting}
reflect(y=8)
 for(i<3)
  if(i>0)
   rectangle(3*i-1,2,3*i,3)
  circle(3*i+1,3*i+1)
\end{lstlisting}
\end{minipage}&$\frac{9}{5} = 1.8\text{x}$ \\\bottomrule
  \end{tabular}
  \end{table}

@@

Given a search policy, what is the best way of using it to quickly find minimum cost programs?
We use a \emph{bias-optimal search algorithm}~\citep{schmidhuber2004optimal}:

@@

Intuitively, an $n$-bias optimal search algorithm spends at least $\probability_{\text{bias}}[\sigma ]/n$ of its time exploring solution space $\sigma $. An example of a $1$-bias optimal search algorithm is a time-sharing system that allocates $\probability_{\text{bias}}[\sigma ]$ of its time to trying $\sigma $. We construct a $1$-bias optimal search algorithm for our domain by identifying $\probability_{\text{bias}}[\sigma ] = \pi_\theta(\sigma |T)$ and $t(\sigma )  $ with how long the synthesizer takes to search  the programs in $\sigma $. This means that the search algorithm explores the entire program space, but spends most of its time searching through the regions of the space that  the policy judges to be most promising.

Now in theory any $\pi_\theta(\cdot |\cdot ) $ is a bias-optimal searcher.
But the actual runtime of the algorithm depends strongly upon
the bias $\probability_{\text{bias}}[\cdot ]$.
Our new approach is to learn $\probability_{\text{bias}}[\cdot ]$
by picking the policy minimizing the
expected bias-optimal time to solve a training corpus, $\mathcal{D}$, of graphics program synthesis problems:
\begin{align}
\textsc{Loss}(\theta ; \mathcal{D})& =  \expect_{T\sim\mathcal{D}}\left[ \min_{\sigma\in \text{\textsc{Best}}(T)}\frac{t(\sigma | T)}{\pi_\theta (\sigma | T)}\right] + \lambda \Vert\theta\Vert_2^2\label{policyLoss}\\
\text{where }  \sigma \in\text{\textsc{Best}}(T) &\text{ if  a minimum cost program for }T \text{ is in }\sigma .\nonumber %\text{cost}(\text{program}(p)) = \min_{p\in \sigma \text{, s.t. }p\text{ evaluates to }T} \text{cost}(p) \nonumber
\end{align}
To generate a training corpus for learning a policy which minimizes this loss,
we synthesized minimum cost programs for each trace of our hand drawings
and for each $\sigma $.
%Solving these
%\footnote{There are 100 the drawings and $2\times 2\times 2\times 3 = 24$ different values of $\sigma $}
%synthesis problems takes a little over a day on 40 CPUs.
%With the training set in hand we  %This loss is differentiable but nonconvex even if $\pi_{\theta}(\cdot |\cdot )^{-1}$ is convex.
We locally minimize this loss using gradient descent. % while annealing a softened minimum to the hard minimization equation~\ref{policyLoss}.
See supplement for details.
Because we want to learn a policy from only $100$ hand-drawn sketches,
we chose a simple low-capacity, bilinear model for a policy:
\begin{equation}
  \pi_{\theta}(\sigma |T)\propto \exp \left( \phi_{\text{params}}(\sigma )^\top\theta \phi_{\text{trace}}(T)\right)
\end{equation}
where $\phi_{\text{params}}(\sigma  )$ is a one-hot encoding of
the parameter settings of $\sigma $ (see Tbl.~\ref{policyOutput})
and $\phi_{\text{trace}}(T)$ extracts a few simple features of the trace $T$;
see supplement for details.
